Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-03 writing/embedded_meta_learning.tex
Integration time
Abstract + intro tweaks

@@

Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrasts, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of architectures which essentially take a functional perspective on meta-learning and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating tasks and data as functions, we can conceptually think of both data \emph{and} learned behaviors as transformable. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing task demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits such as a useful framework for continual learning. \par
\section{Method}
\subsection{Architecture}
The essential idea of the architecture is that we separate the system into two parts: 

@@

The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior). This is a step closer to the flexible reasoning humans display. \par

@@

\textbf{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out task from a few examples in Fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``Losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. It is worth noting that the sample complexity in terms of number of tasks trained on is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}.\par

@@

\includegraphics[width=0.5\textwidth]{figures/basic_meta_learning.png}

@@

>>>>

\begin{abstract}
How can deep learning systems flexibly reuse their knowledge? We suggest that the key is representing the task being performed along with the computations used to perform it. Specifically, we propose a new class of challenges, meta-mappings, which involve systematically transforming task behaviors. We draw inspiration from meta-learning and functional programming to propose a class of architectures that represent both data and tasks in a shared latent space. We demonstrate the flexibility of these architectures by showing that they can exhibit zero-shot remapping of behavior on new tasks. 
\end{abstract}

@@

Humans are able to use and reuse knowledge more flexibly than most deep learning models can \citep[e.g.][]{Lake2016}. One fundamental reason for this is that humans are aware of what we are trying to compute and why. By contrast, there is a fundamental separation of knowledge within most deep neural networks -- although deep networks represent knowledge about data (in their activations) and knowledge about the structure of tasks (in their parameters), they do not represent any relationships between data and tasks. That is, neural networks knowledge about what is being computed is only implicitly accessible to those computations. \par
There are a number of advantages to representing knowledge about data and tasks together. In particular, it can grant the ability to rapidly adapt behavior to a new task. The problem of rapid learning has been partially addressed by meta-learning systems \citep[e.g.][]{Vinyals2016, Finn2017a}. However, humans can take our knowledge and flexibly alter our behavior in accordance with a change in task demands or a single instruction. For example, once we learn to play a game, we can immediately switch to playing in order to lose, and can achieve reasonable performance without any retraining (i.e. zero-shot). Deep learning systems at present generally lack this representational flexiblity. \par
In this paper, we propose a new class of tasks based on this idea: meta-mappings, i.e. mappings between tasks. For example, once a system has been trained to play a game, it should be able to alter its strategy in accordance with a change in task demands, such as trying to lose the game, or trying to play with an alteration in the reward function. The challenge is that the system must adapt to performing these tasks without seeing any data from the new task, only from either examples of similar transformations on different tasks (``here's how to change your go strategy to a losing one, here's how to change your checkers strategy to a losing one, now try to lose at chess'') or from natural language instructions (``try to lose at chess''). These mappings essentially provide a new opportunity for zero-shot transfer from one task to another -- that is, the system must transform its behavior to adapt to a new task without ever having seen training data from that task. This type of transfer is easily accessible to humans \citep{Lake2016}, but is generally inaccessible to most deep-learning models. \par
To address this challenge, we propose a new class of architectures which essentially take a functional perspective on meta-learning, and exploit the idea of homoiconicity\footnote{A homoiconic programming language is one in which programs in the language can be manipulated by programs in the language, just as data can.}. By treating task behaviors and data as functions, we can conceptually think of both data \emph{and} learned task behaviors as transformable. By embedding task behaviors, data, and language in the same space, we can conceptually think of both basic task solutions and adaptive functions, such as following instructions or altering behavior, as the same types of functional mappings. This yields the ability to not only learn to solve new tasks, but to learn how to flexibly transform these solutions in response to changing demands. We suggest that this ability will be key to building more intelligent and flexible deep learning systems. \par
In particular, we propose a meta-learning architecture which embeds data (including natural language inputs) and tasks in a shared space, and meta-learns to execute meta-mappings over this space. We demonstrate that this approach allows flexible remapping of behavior, as well as offering other benefits, such as a useful framework for continual learning. \par
\section{Embedded Meta-Learning (EML) architecture}
We propose a new class of Embedded Meta-Learning (EML) architectures which essentially separate the system into two parts: 

@@

The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior based on examples). This is a step closer to the flexible reasoning humans display. \par

@@

\textbf{Basic meta learning:} First, we show that the system is able to achieve the basic goal of learning a held-out task from a few examples in Fig. \ref{cards_basic_results}. We compare two different possible hold-out sets: 1) train on half the tasks at random, or 2) specifically hold out all the ``Losers'' variations of the ``straight flush'' game. In either of these cases, the meta-learning system achieves well above chance performance (0) at the held out tasks, although it is slightly worse at generalizing to the targeted hold out, despite having more training tasks in that case. It is worth noting that the sample complexity in terms of number of tasks trained on is not that high, even training on 20 randomly selected tasks leads to good generalization to the held-out tasks. Furthermore, the task embeddings generated by the system are organized by the features of the games, see appendix \ref{app_cards_tsne}. Finally, the system can integrate new tasks quite rapidly (fig. \ref{cards_basic_integration_results}).\par

@@

\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning.png}

@@

\begin{subfigure}[t]{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/basic_meta_learning_integrating_over_time.png}
\caption{The system can rapidly adapt to new tasks.}
\label{cards_basic_integration_results}
